---
title: "The interactive origin of iconiciy: Permutation tests"
output: pdf_document
---

# Load libraries

```{r warning=FALSE, message=FALSE}

library(gplots)
library(lattice)
library(lme4)
```

```{r echo=FALSE}
setwd("~/Documents/MPI/MonicaIconicity/SelectionAnalysis/analysis/")
```

# Load data

```{r}
finalLangs = read.csv("../data/finalLanguages/FinalLanguages.csv", stringsAsFactors = F)
# make a variable which stores condition, chain and generation
finalLangs$cond2 = paste(finalLangs$Cond,finalLangs$Chain, finalLangs$Gen)
```

# Run permutation test

For each output language, take the difference in means in spikiness ratings between spiky and non-spiky meanings.  Compare this to 1000 permutations of the numbers.

```{r perm1, cache=TRUE}

# Set the random seed for reproducibility
set.seed(1278)

# what factor should the data be split by?
split = finalLangs[finalLangs$cond2==unique(finalLangs$cond2)[1],]$Shape

# for each language (a single generation's output)
res = tapply(finalLangs$RatedSpikiness, finalLangs$cond2, function(X){
  # calculate the true difference
  trueDiff = -diff(tapply(X, split,mean))
  # permute the numbers and re-calculate difference
  permDiff = replicate(1000,
        {-diff(tapply(sample(X), split,mean)) })
  
  # work out p and z-socres
  p = 1- sum(trueDiff > permDiff ) / length(permDiff)
  z.score = (trueDiff - mean(permDiff)) / sd(permDiff)
  return(c(p,z.score))
})

```

Recast results into data frame:

```{r}

res2 = data.frame(
p = sapply(res,head,n=1),
z = sapply(res,tail,n=1),
t(sapply(names(res),function(X){
  strsplit(X," ")[[1]]
}))
, stringsAsFactors = F)
names(res2) = c("p",'z','condition','chain','gen')
```

Plot the results.  Each line represents an independent chain.  The red lines show the results for the Learning condition.  The results suggest that there is no strong difference between the conditions.  One of the learning chains decreases in iconicity, due to that chain focussing on specifying the colour rather than the shape.

```{r}
par(mfrow=c(1,2))
plot(c(1,6), c(-3,3), type='n',ylab='Z score', xlab='Generations')
for(i in unique(res2$chain)){
  dx =res2[res2$chain==i,]
  lines(dx$gen,dx$z, col=c("black","red")[(dx$condition=="Learn")+1])
}

plot(c(1,6), c(0,1), type='n',ylab='p', xlab='Generations')
for(i in unique(res2$chain)){
  dx =res2[res2$chain==i,]
  lines(dx$gen,dx$p, col=c("black","red")[(dx$condition=="Learn")+1])
}
```

```{r}
rownames(res2) = NULL
res2
```

\newpage

# Looking only at red-coloured meanings

The mixed effects results suggested that there is a difference between conditions.  The difference could be because the permutation test does not take into account effects of items nor the other aspects of meaning.

```{r cache=TRUE}
finalLangs2 = finalLangs[finalLangs$Colour=="Rojo",]


# Set the random seed for reproducibility
set.seed(1278)

# what factor should the data be split by?
split = finalLangs2[finalLangs2$cond2==unique(finalLangs2$cond2)[1],]$Shape

# for each language (a single generation's output)
resRed = tapply(finalLangs2$RatedSpikiness, finalLangs2$cond2, function(X){
  # calculate the true difference
  trueDiff = -diff(tapply(X, split,mean))
  # permute the numbers and re-calculate difference
  permDiff = replicate(1000,
        {-diff(tapply(sample(X), split,mean)) })
  
  # work out p and z-socres
  p = 1- sum(trueDiff > permDiff ) / length(permDiff)
  z.score = (trueDiff - mean(permDiff)) / sd(permDiff)
  return(c(p,z.score))
})
res2Red = data.frame(
p = sapply(resRed,head,n=1),
z = sapply(resRed,tail,n=1),
t(sapply(names(resRed),function(X){
  strsplit(X," ")[[1]]
}))
, stringsAsFactors = F)
names(res2Red) = c("p",'z','condition','chain','gen')

```

Plot the results.  In this case, we do see a division between the two conditions by the last generation.

```{r}
par(mfrow=c(1,2))
plot(c(1,6), c(-3,3), type='n',ylab='Z score')
for(i in unique(res2Red$chain)){
  dx =res2Red[res2Red$chain==i,]
  lines(dx$gen,dx$z, col=c("black","red")[(dx$condition=="Learn")+1])
}

plot(c(1,6), c(0,1), type='n',ylab='p')
for(i in unique(res2Red$chain)){
  dx =res2Red[res2Red$chain==i,]
  lines(dx$gen,dx$p, col=c("black","red")[(dx$condition=="Learn")+1])
}
```

\newpage

# Stratified permutation

More generally, the results change if we run a permutation test on the whole data, but only permuting within colour.

```{r perm2, cache=TRUE}
set.seed(2389)
split = finalLangs[finalLangs$cond2==unique(finalLangs$cond2)[1],]$Shape
permWithin = finalLangs[finalLangs$cond2==unique(finalLangs$cond2)[1],]$Colour


res3 = tapply(finalLangs$RatedSpikiness, finalLangs$cond2, function(X){
  trueDiff = -diff(tapply(X, split,mean))
  permDiff = replicate(1000,
                       {-diff(tapply(
                         unlist(tapply(X,permWithin,sample)), split,mean)) })
  p = 1- sum(trueDiff > permDiff ) / length(permDiff)
  z.score = (trueDiff - mean(permDiff)) / sd(permDiff)
  return(c(p,z.score))
})

res4 = data.frame(
  p = sapply(res3,head,n=1),
  z = sapply(res3,tail,n=1),
  t(sapply(names(res3),function(X){
    strsplit(X," ")[[1]]
  }))
  , stringsAsFactors = F)
names(res4) = c("p",'z','condition','chain','gen')
```

```{r}

par(mfrow=c(1,2))
plot(c(1,6), c(-20,20), type='n',ylab='Z score',xlab='Generations')
for(i in unique(res4$chain)){
  dx =res4[res4$chain==i,]
  lines(dx$gen,dx$z, col=c("black","red")[(dx$condition=="Learn")+1])
}

plot(c(1,6), c(0,1), type='n',ylab='p',xlab='Generations')
for(i in unique(res4$chain)){
  dx =res4[res4$chain==i,]
  lines(dx$gen,dx$p, col=c("black","red")[(dx$condition=="Learn")+1])
}
```


```{r}
rownames(res4) = NULL
res4
```
